# (PART\*) Getting started {-}

# Your first Shiny app 

## Single file app

```{r eval = FALSE}
library(shiny)

ui <- fluidPage(
  h1("Hello World!")
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


## Adding UI controls

```{r eval = FALSE}
library(shiny)

ui <- fluidPage(
  selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)

server <- function(input, output, session) {
  output$summary <- renderPrint({
    dataset <- get(input$dataset, "package:datasets")
    summary(dataset)
  })
  
  output$table <- renderTable({
    dataset <- get(input$dataset, "package:datasets")
    dataset
  })
}

shinyApp(ui, server)
```

## Reactive expression

You create a reactive expression by wrapping a block of code in reactive({...}) and assigning it to a variable, and you use a reactive expression by calling it like a function. But while it looks like you’re **calling a function**, a reactive expression has an important difference: it only runs the first time it is called and then it **caches** its result until it needs to be updated.

```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
  selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)

server <- function(input, output, session) {
  # Create a reactive expression
  dataset <- reactive({
    get(input$dataset, "package:datasets")
  })

  output$summary <- renderPrint({
    # Use a reactive expression by calling it like a function
    summary(dataset())
  })
  
  output$table <- renderTable({
    dataset()
  })
}

shinyApp(ui, server)
```

## Exercises

### Exercise 1

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  textInput("name", "What's your name?"),
  textOutput("greeting")
)

server <- function(input, output, session) {
  output$greeting <- renderText({
  paste0("Hello ", input$name)
})
}

shinyApp(ui, server)
```

### Exercise 2

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  "then x times 5 is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * 5
  })
}

shinyApp(ui, server)
```

### Exercise 3

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", label = "and y is", min = 1, max = 50, value = 5),
  "then, x times y is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * input$y
  })
}

shinyApp(ui, server)
```

### Exercise 4

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  sliderInput("x", "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", "and y is", min = 1, max = 50, value = 5),
  "then, (x * y) is", textOutput("product"),
  "and, (x * y) + 5 is", textOutput("product_plus5"),
  "and (x * y) + 10 is", textOutput("product_plus10")
)

server <- function(input, output, session) {

  product <- reactive({
    input$x * input$y
  })
  output$product <- renderText({ 
    product()
  })
  output$product_plus5 <- renderText({ 
    product() + 5
  })
  output$product_plus10 <- renderText({ 
    product() + 10
  })
}

shinyApp(ui, server)
```

### Exercise 5

```{r eval=FALSE}
library(shiny)
library(ggplot2)

datasets <- c("economics", "faithfuld", "seals")
ui <- fluidPage(
  selectInput("dataset", "Dataset", choices = datasets),
  verbatimTextOutput("summary"),
  plotOutput("plot")
)

server <- function(input, output, session) {
  dataset <- reactive({
    get(input$dataset, "package:ggplot2")
  })
  output$summary <- renderPrint({
    summary(dataset())
  })
  output$plot <- renderPlot({
    plot(dataset())
  }, res = 96)
}

shinyApp(ui, server)
```

