# Basic UI

## Shiny extensions

+ shinyWidgets
+ colorpicker
+ sorttable
+ <https://github.com/nanxstats/awesome-shiny-extensions>

## Inputs

### Common structure

All input functions have the same first argument: `inputId`.

The inputId has two constraints:

+ It must be a simple string that contains only letters, numbers, and underscores (no spaces, dashes, periods, or other special characters allowed!). Name it like you would name a variable in R.

+ It must be unique. If it’s not unique, you’ll have no way to refer to this control in your server function!

### Input examples

```{r, eval=FALSE}
library(shiny)

animals <- c("dog", "cat", "mouse", "bird", "other", "I hate animals")

ui <- fluidPage(
  # free text
  textInput("name", "What's your name?"),
  passwordInput("password", "What's your password?"),
  textAreaInput("story", "Tell me about yourself", rows = 3),

  # numeric inputs
  numericInput("num", "Number one", value = 0, min = 0, max = 100),
  sliderInput("num2", "Number two", value = 50, min = 0, max = 100),
  sliderInput("rng", "Range", value = c(10, 20), min = 0, max = 100),

  # date inputs
  dateInput("dob", "When were you born?"),
  dateRangeInput("holiday", "When do you want to go on vacation next?"),

  # limited choices
  selectInput("state", "What's your favourite state?", state.name),
  radioButtons("animal", "What's your favourite animal?", animals),
  # we can even use Icon
  radioButtons("rb", "Choose one:",
    choiceNames = list(
      icon("angry"),
      icon("smile"),
      icon("sad-tear")
    ),
    choiceValues = list("angry", "happy", "sad")
  ),

  # select input
  selectInput(
    "state", "What's your favourite state?", state.name,
    multiple = TRUE
  ),

  # file upload
  fileInput("upload", NULL),

  # action button
  actionButton("click", "Click me!", class = "btn-danger"),
  actionButton("drink", "Drink me!", icon = icon("cocktail"), class = "btn-lg btn-success")


)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

### Exercises

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  # ex 1
  textInput('t1', label = NULL, placeholder = 'Your name'),

  # ex 2
  sliderInput(
    's1', 
    label = 'When should we deliver?', 
    min = as.Date('2020-09-16'), 
    max = as.Date('2020-09-23'), 
    value = as.Date('2020-09-17')
  ),

  # ex 3
  sliderInput(
    's2',
    'Ex 3',
    min = 0,
    max = 100,
    value = 50,
    step = 5,
    animate = TRUE
  ),

  # ex 4
  selectInput(
    'se1',
    'Select an option',
    choices = list(
      group1 = list('Option 1', 'Option 2'),
      group2 = list('Option 3', 'Option 4', 'Option 5')
    )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Outputs

Outputs in the UI **create placeholders** that are later filled by the server function.

### Text output

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  textOutput("text"),
  verbatimTextOutput("code")
)
server <- function(input, output, session) {
  output$text <- renderText({ 
    "Hello friend!" 
  })
  output$code <- renderPrint({ 
    summary(1:10) 
  })
}

shinyApp(ui, server)
```

### Table output

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  tableOutput("static"),
  dataTableOutput("dynamic")
)
server <- function(input, output, session) {
  output$static <- renderTable(head(mtcars))
  output$dynamic <- renderDataTable(mtcars, options = list(pageLength = 5))
}

shinyApp(ui, server)
```

### Plot output

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  plotOutput("plot", width = "400px")
)
server <- function(input, output, session) {
  output$plot <- renderPlot(plot(1:5), res = 96)
}

shinyApp(ui, server)
```

Plots are special because they are outputs that **can also act as inputs**. 

plotOutput() has a number of arguments like `click`, `dblclick`, and `hover`. If you pass these a string, like click = "plot_click", they’ll create a reactive input (input$plot_click) that you can use to handle user interaction on the plot.

### Exercises

#### Ex 1

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  textOutput('t1'),
  verbatimTextOutput('v1'),
  verbatimTextOutput('v2'),
  verbatimTextOutput('v3')
)

server <- function(input, output, session) {
  output$t1 <- renderText("Good morning!")
  output$v1 <- renderPrint(str(lm(mpg ~ wt, data = mtcars)))
  output$v2 <- renderPrint(summary(mtcars))
  output$v3 <- renderPrint(t.test(1:5, 2:6))
}

shinyApp(ui, server)
```

#### Ex 2

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  plotOutput("plot", width = "400px")
)
server <- function(input, output, session) {
  output$plot <- renderPlot(plot(1:5), 
    width = 700, 
    height = 300, 
    res = 96, 
    alt = 'a scatterplot of five random numbers')
}

shinyApp(ui, server)
```

#### Ex 3

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  dataTableOutput("table")
)

server <- function(input, output, session) {
  output$table <- renderDataTable(mtcars, options = list(dom='t'))
}

shinyApp(ui, server)
```

#### Ex 4

```{r, eval=FALSE}
library(shiny)
library(reactable)

ui <- fluidPage(
  reactableOutput("table")
)

server <- function(input, output) {
  output$table <- renderReactable({
    reactable(mtcars)
  })
}

shinyApp(ui, server)
```

