# Basic reactivity

The key idea of reactive programming is to specify **a graph of dependencies** so that when an input changes, all related outputs are automatically updated. 

## Server function

The ui is simple because every user gets the same HTML. The server is more complicated because every user needs to get an independent version of the app; when user A moves a slider, user B shouldn’t see their outputs change.

To achieve this independence, Shiny invokes your `server()` function each time a new session starts.

### Input

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  numericInput("count", label = "Number of values", value = 100)
)

server <- function(input, output, session) {
  # input objects are read-only
  input$count <- 10  
}

shinyApp(ui, server)
```

`input` is selective about who is allowed to read it. To read from an input, you must be in a reactive context created by a function like `renderText()` or `reactive()`

### Output

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  textOutput("greeting")
)

server <- function(input, output, session) {
  output$greeting <- renderText("Hello human!")
}

shinyApp(ui, server)
```

You always use the output object in concert with a render function.

The `render` function does two things:

+ It sets up a special reactive context that automatically tracks what inputs the output uses.

+ It converts the output of your R code into HTML suitable for display on a web page.


## Reactive programming

### Reactive programming exercises

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  textInput("name", "What's your name?"),
  textOutput("greeting")
)

# server1
# server <- function(input, output, server) {
#   output$greeting <- renderText(paste0("Hello ", input$name))
# }

# server2
# server <- function(input, output, server) {
#   greeting <- reactive(paste0("Hello ", input$name))
#   output$greeting <- renderText(greeting())
# }

# server3
server <- function(input, output, server) {
  output$greeting <- renderText(paste0("Hello ", input$name))
}

shinyApp(ui, server)
```

## Reactive Expressions

```{r, eval=FALSE}
library(ggplot2)

freqpoly <- function(x1, x2, binwidth = 0.1, xlim = c(-3, 3)) {
  df <- data.frame(
    x = c(x1, x2),
    g = c(rep("x1", length(x1)), rep("x2", length(x2)))
  )

  ggplot(df, aes(x, colour = g)) +
    geom_freqpoly(binwidth = binwidth, size = 1) +
    coord_cartesian(xlim = xlim)
}

t_test <- function(x1, x2) {
  test <- t.test(x1, x2)
  
  # use sprintf() to format t.test() results compactly
  sprintf(
    "p value: %0.3f\n[%0.2f, %0.2f]",
    test$p.value, test$conf.int[1], test$conf.int[2]
  )
}

# compare two simulated datasets with a plot and a hypothesis test
set.seed(100)
x1 <- rnorm(100, mean = 0, sd = 0.5)
x2 <- rnorm(200, mean = 0.15, sd = 0.9)

freqpoly(x1, x2)
t_test(x1, x2)

```


```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  fluidRow(
    column(4, 
      "Distribution 1",
      numericInput("n1", label = "n", value = 1000, min = 1),
      numericInput("mean1", label = "µ", value = 0, step = 0.1),
      numericInput("sd1", label = "σ", value = 0.5, min = 0.1, step = 0.1)
    ),
    column(4, 
      "Distribution 2",
      numericInput("n2", label = "n", value = 1000, min = 1),
      numericInput("mean2", label = "µ", value = 0, step = 0.1),
      numericInput("sd2", label = "σ", value = 0.5, min = 0.1, step = 0.1)
    ),
    column(4,
      "Frequency polygon",
      numericInput("binwidth", label = "Bin width", value = 0.1, step = 0.1),
      sliderInput("range", label = "range", value = c(-3, 3), min = -5, max = 5)
    )
  ),
  fluidRow(
    column(9, plotOutput("hist")),
    column(3, verbatimTextOutput("ttest"))
  )
)

server <- function(input, output, session) {
  x1 <- reactive(rnorm(input$n1, input$mean1, input$sd1))
  x2 <- reactive(rnorm(input$n2, input$mean2, input$sd2))

  output$hist <- renderPlot({
    freqpoly(x1(), x2(), binwidth = input$binwidth, xlim = input$range)
  }, res = 96)

  output$ttest <- renderText({
    t_test(x1(), x2())
  })
}

shinyApp(ui, server)
```

## Controlling timeing of evaluation

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  fluidRow(
    column(3, 
      numericInput("lambda1", label = "lambda1", value = 3),
      numericInput("lambda2", label = "lambda2", value = 5),
      numericInput("n", label = "n", value = 1e4, min = 0)
    ),
    column(9, plotOutput("hist"))
  )
)

server <- function(input, output, session) {
  timer <- reactiveTimer(500) # the plot will update twice a second
  
  x1 <- reactive({
    timer()
    rpois(input$n, input$lambda1)
  })
  
  x2 <- reactive({
    timer()
    rpois(input$n, input$lambda2)
  })
  
  output$hist <- renderPlot({
    freqpoly(x1(), x2(), binwidth = 1, xlim = c(0, 40))
  }, res = 96)
}

shinyApp(ui, server)
```

### eventReactive

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  fluidRow(
    column(3, 
      numericInput("lambda1", label = "lambda1", value = 3),
      numericInput("lambda2", label = "lambda2", value = 5),
      numericInput("n", label = "n", value = 1e4, min = 0),
      actionButton("simulate", "Simulate!")
    ),
    column(9, plotOutput("hist"))
  )
)

server <- function(input, output, session) {
  x1 <- eventReactive(input$simulate, {
    rpois(input$n, input$lambda1)
  })
  x2 <- eventReactive(input$simulate, {
    rpois(input$n, input$lambda2)
  })

  output$hist <- renderPlot({
    freqpoly(x1(), x2(), binwidth = 1, xlim = c(0, 40))
  }, res = 96)
}

shinyApp(ui, server)
```

## Observers

Sometimes you need to reach outside of the app and cause side-effects to happen elsewhere in the world. This might be saving a file to a shared network drive, sending data to a web API, updating a database, or (most commonly) printing a debugging message to the console. These actions don’t affect how your app looks, so you shouldn’t use an output and a render function. Instead you need to use an observer.

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  textInput("name", "What's your name?"),
  textOutput("greeting")
)

server <- function(input, output, session) {
  string <- reactive(paste0("Hello ", input$name, "!"))
  
  output$greeting <- renderText(string())
  observeEvent(input$name, {
    message("Greeting performed")
  })
}

shinyApp(ui, server)
```

