# Dynamic UI

:::{.rmdnote}

There are 3 key techniques for creating dynamic user interfaces:

+ Using the `update` family of functions to modify parameters of input controls

+ Using `tabsetPanel()` to conditionally show and hide parts of the user interface

+ Using the `uiOutput()` and `renderUI()` to generate selected parts of the user interface with code

:::

## Updating inputs

```{r,eval=FALSE}
library(shiny)

ui <- fluidPage(
  numericInput("min", "Minimum", 0),
  numericInput("max", "Maximum", 3),
  sliderInput("n", "n", min = 0, max = 3, value = 1)
)

server <- function(input, output, session) {
  observeEvent(input$min, {
    updateSliderInput(inputId = "n", min = input$min)
  })  
  observeEvent(input$max, {
    updateSliderInput(inputId = "n", max = input$max)
  })
}

shinyApp(ui, server)

```

:::{.rmdnote}
The `updateXXX()` functions: take the name of the input (as a string) as the `inputId` argument. 
:::

### reset button

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  sliderInput("x1", "x1", 0, min = -10, max = 10),
  sliderInput("x2", "x2", 0, min = -10, max = 10),
  sliderInput("x3", "x3", 0, min = -10, max = 10),
  actionButton("reset", "Reset")
)

server <- function(input, output, session) {
  observeEvent(input$reset, {
    updateSliderInput(inputId = "x1", value = 0)
    updateSliderInput(inputId = "x2", value = 0)
    updateSliderInput(inputId = "x3", value = 0)
  })
}

shinyApp(ui, server)
```

### update button label

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  numericInput("n", "Simulations", 10),
  actionButton("simulate", "Simulate")
)

server <- function(input, output, session) {
  observeEvent(input$n, {
    label <- paste0("Simulate ", input$n, " times")
    updateActionButton(inputId = "simulate", label = label)
  })
}

shinyApp(ui, server)
```

### hierarchical select boxes
```{r, eval=FALSE}
library(vroom)
library(dplyr)
library(shiny)

sales <- vroom::vroom("assets/sales_data_sample.csv", col_types = list(), na = "")

ui <- fluidPage(
  selectInput("territory", "Territory", choices = unique(sales$TERRITORY)),
  selectInput("customername", "Customer", choices = NULL),
  selectInput("ordernumber", "Order number", choices = NULL),
  tableOutput("data")
)

server <- function(input, output, session) {
  territory <- reactive({
    filter(sales, TERRITORY == input$territory)
  })
  observeEvent(territory(), {
    choices <- unique(territory()$CUSTOMERNAME)
    updateSelectInput(inputId = "customername", choices = choices) 
  })
  
  customer <- reactive({
    req(input$customername)
    filter(territory(), CUSTOMERNAME == input$customername)
  })
  observeEvent(customer(), {
    choices <- unique(customer()$ORDERNUMBER)
    updateSelectInput(inputId = "ordernumber", choices = choices)
  })
  
  output$data <- renderTable({
    req(input$ordernumber)
    customer() %>% 
      filter(ORDERNUMBER == input$ordernumber) %>% 
      select(QUANTITYORDERED, PRICEEACH, PRODUCTCODE)
  })
}

shinyApp(ui, server)

```


### `freezeReactiveValue()`

before using `freezeReactiveValue`

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  selectInput("dataset", "Choose a dataset", c("pressure", "cars")),
  selectInput("column", "Choose column", character(0)),
  verbatimTextOutput("summary")
)

server <- function(input, output, session) {
  dataset <- reactive(get(input$dataset, "package:datasets"))
  
  observeEvent(input$dataset, {
    updateSelectInput(inputId = "column", choices = names(dataset()))
  })
  
  output$summary <- renderPrint({
    summary(dataset()[[input$column]])
  })
}

shinyApp(ui, server)
```

after using `freezeReactiveValue`

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  selectInput("dataset", "Choose a dataset", c("pressure", "cars")),
  selectInput("column", "Choose column", character(0)),
  verbatimTextOutput("summary")
)

server <- function(input, output, session) {
  dataset <- reactive(get(input$dataset, "package:datasets"))
  
  observeEvent(input$dataset, {
    freezeReactiveValue(input, "column")
    updateSelectInput(inputId = "column", choices = names(dataset()))
  })
  
  output$summary <- renderPrint({
    summary(dataset()[[input$column]])
  })
}

shinyApp(ui, server)
```

:::{.rmdnote}
`freezeReactiveValue()` ensures that any reactives or outputs that use the input won't be 
updated until the next full round of invalidation
:::


### inter-related inputs

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  numericInput("temp_c", "Celsius", NA, step = 1),
  numericInput("temp_f", "Fahrenheit", NA, step = 1)
)

server <- function(input, output, session) {
  observeEvent(input$temp_f, {
    c <- round((input$temp_f - 32) * 5 / 9)
    updateNumericInput(inputId = "temp_c", value = c)
  })
  
  observeEvent(input$temp_c, {
    f <- round((input$temp_c * 9 / 5) + 32)
    updateNumericInput(inputId = "temp_f", value = f)
  })
}

shinyApp(ui, server)
```

### Exercises

#### Ex 1

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  numericInput("year", "year", value = 2020),
  dateInput("date", "date")
)

server <- function(input, output, session) {
  observeEvent(input$year, {
    updateDateInput(inputId = "date", value =paste0(input$year, '-01-01'), min = paste0(input$year,'-01-01'), max = paste0(input$year, '-12-31'))
  })
}

shinyApp(ui, server)
```

#### Ex 2

```{r, eval=FALSE}
library(shiny)
library(openintro)

states <- unique(county$state)

ui <- fluidPage(
  selectInput("state", "State", choices = states),
  selectInput("county", "County", choices = NULL)
)

server <- function(input, output, session) {
  observeEvent(input$state, {
    freezeReactiveValue(input, 'county')
    county_choices <- unique(county[county$state=='Alabama',][['name']])
    if (input$state %in% c('Louisiana', 'Alaska')) lbl <- 'Parish' else lbl <- 'County'
    updateSelectInput(inputId = 'county', label = lbl, choices = county_choices )
  })
}

shinyApp(ui, server)
```

#### Ex 3

```{r, eval=FALSE}
library(shiny)
library(gapminder)

continents <- unique(gapminder$continent)

ui <- fluidPage(
  selectInput("continent", "Continent", choices = continents), 
  selectInput("country", "Country", choices = NULL),
  tableOutput("data")
)

server <- function(input, output, session) {
  observeEvent(input$continent, {
    freezeReactiveValue(input, 'country')
    updateSelectInput(inputId = 'country', choices = unique(gapminder[gapminder$continent==input$continent,'country', drop=TRUE]))
  })
  
  output$data <- renderTable({
    gapminder[gapminder$country==input$country,]
  })
}

shinyApp(ui, server)
```

#### Ex 4

```{r, eval=FALSE}
library(shiny)
library(gapminder)

continents <- as.character(unique(gapminder$continent))

ui <- fluidPage(
  selectInput("continent", "Continent", choices = c('All', continents)), 
  selectInput("country", "Country", choices = NULL),
  tableOutput("data")
)

server <- function(input, output, session) {
  observeEvent(input$continent, {
    freezeReactiveValue(input, 'country')
    if (input$continent == 'All') {
      countries <- unique(gapminder[['country']])
      countries <- c('All', as.character(countries))
    } else countries <- unique(gapminder[gapminder$continent==input$continent,'country', drop=TRUE])
    updateSelectInput(inputId = 'country', choices = countries)
  })
  
  output$data <- renderTable({
    if (input$country == 'All') {
      gapminder
    } else gapminder[gapminder$country==input$country,]
  })
}

shinyApp(ui, server)
```

#### Ex 5

circular references


## Dynamic visibility

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput("controller", "Show", choices = paste0("panel", 1:3))
    ),
    mainPanel(
      tabsetPanel(
        id = "switcher",
        type = "hidden",
        tabPanelBody("panel1", "Panel 1 content"),
        tabPanelBody("panel2", "Panel 2 content"),
        tabPanelBody("panel3", "Panel 3 content")
      )
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$controller, {
    updateTabsetPanel(inputId = "switcher", selected = input$controller)
  })
}

shinyApp(ui, server)

```

### Conditional UI

```{r, eval=FALSE}
library(shiny)

parameter_tabs <- tabsetPanel(
  id = "params",
  type = "hidden",
  tabPanel("normal",
    numericInput("mean", "mean", value = 1),
    numericInput("sd", "standard deviation", min = 0, value = 1)
  ),
  tabPanel("uniform", 
    numericInput("min", "min", value = 0),
    numericInput("max", "max", value = 1)
  ),
  tabPanel("exponential",
    numericInput("rate", "rate", value = 1, min = 0),
  )
)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput("dist", "Distribution", 
        choices = c("normal", "uniform", "exponential")
      ),
      numericInput("n", "Number of samples", value = 100),
      parameter_tabs,
    ),
    mainPanel(
      plotOutput("hist")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$dist, {
    updateTabsetPanel(inputId = "params", selected = input$dist)
  }) 
  
  sample <- reactive({
    switch(input$dist,
      normal = rnorm(input$n, input$mean, input$sd),
      uniform = runif(input$n, input$min, input$max),
      exponential = rexp(input$n, input$rate)
    )
  })
  output$hist <- renderPlot(hist(sample()), res = 96)
}

shinyApp(ui, server)
```


### Wizard interface

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  tabsetPanel(
    id = "wizard",
    type = "hidden",
    tabPanel("page_1", 
      "Welcome!",
      actionButton("page_12", "next")
    ),
    tabPanel("page_2", 
      "Only one page to go",
      actionButton("page_21", "prev"),
      actionButton("page_23", "next")
    ),
    tabPanel("page_3", 
      "You're done!",
      actionButton("page_32", "prev")
    )
  )
)

server <- function(input, output, session) {
  switch_page <- function(i) {
    updateTabsetPanel(inputId = "wizard", selected = paste0("page_", i))
  }
  
  observeEvent(input$page_12, switch_page(2))
  observeEvent(input$page_21, switch_page(1))
  observeEvent(input$page_23, switch_page(3))
  observeEvent(input$page_32, switch_page(2))
}

shinyApp(ui, server)
```

### Exercises

#### Ex 1

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  tabsetPanel(
    tabPanel(
      'User Selection',
      selectInput('s1', 'Select one', choices = c('a','b','c')),
      checkboxInput('c1', 'advanced')
      # actionButton('b1', 'click me')
    ),
    tabPanel(
      title = 'Tab 2',
      h1('tab 2')
    ),
    id = 'tabs',
    type = 'hidden'
  )
)

server <- function(input, output, session) {
  observeEvent(input$c1, {
    if (input$c1) updateTabsetPanel(inputId = 'tabs', selected = 'Tab 2')
  })
  
  # observe(print(input$s1))
}

shinyApp(ui, server)
```

#### Ex 2

```{r, eval=FALSE}
library(shiny)
library(ggplot2)

plot_tabs <- tabsetPanel(
  id = "plots",
  type = "hidden",
  tabPanel("histogram",
    numericInput("hbinw", "binwidth", value = 0.1, min=0, max=1, step = 0.1)
  ),
  tabPanel("freqpoly", 
    numericInput("fbinw", "binwidth", value = 0.1, min=0, max=1, step = 0.1)
  ),
  tabPanel("density",
    numericInput("bw", "bw", value = 0.1, min=0, max=1, step = 0.1)
  )
)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput("type", "Plot Type", 
        choices = c("histogram", "freqpoly", "density")
      ),
      plot_tabs,
    ),
    mainPanel(
      plotOutput("plot")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$type, {
    updateTabsetPanel(inputId = "plots", selected = input$type)
  }) 
  
  output$plot <- renderPlot({
    p <- ggplot(diamonds, aes(carat))

    if (input$type == 'histogram') p + geom_histogram(binwidth = input$hbinw)
    else if (input$type == 'freqpoly') p + geom_freqpoly(binwidth = input$fbinw)
    else if (input$type == 'density') p + geom_density(bw = input$bw)
  })
}

shinyApp(ui, server)
```

#### Ex 3

```{r, eval=FALSE}
library(shiny)
library(ggplot2)

plot_tabs <- tabsetPanel(
  id = "plots",
  type = "hidden",
  tabPanel("histogram",
    numericInput("hbinw", "binwidth", value = 0.1, min=0, max=1, step = 0.1)
  ),
  tabPanel("freqpoly", 
    numericInput("fbinw", "binwidth", value = 0.1, min=0, max=1, step = 0.1)
  ),
  tabPanel("density",
    numericInput("bw", "bw", value = 0.1, min=0, max=1, step = 0.1)
  )
)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput("type", "Plot Type", 
        choices = c("histogram", "freqpoly", "density")
      ),
      plot_tabs,
      checkboxGroupInput('cb','Show plot(s)', choices = c("histogram", "freqpoly", "density"))
    ),
    mainPanel(
      plotOutput("plot")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$type, {
    updateTabsetPanel(inputId = "plots", selected = input$type)
  }) 
  
  output$plot <- renderPlot({
    p <- ggplot(diamonds, aes(carat))

    if ('histogram' %in% input$cb) p <- p + geom_histogram(binwidth = input$hbinw)
    if ('freqpoly' %in% input$cb) p <- p + geom_freqpoly(binwidth = input$fbinw)
    if ('density' %in% input$cb) p <- p + geom_density(bw = input$bw)
    
    p
  })

}

shinyApp(ui, server)
```

## Creating UI with code

### Getting started

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  textInput("label", "label"),
  selectInput("type", "type", c("slider", "numeric")),
  uiOutput("numeric")
)
server <- function(input, output, session) {
  output$numeric <- renderUI({
    if (input$type == "slider") {
      sliderInput("dynamic", input$label, value = 0, min = 0, max = 10)
    } else {
      numericInput("dynamic", input$label, value = 0, min = 0, max = 10) 
    }
  })
}

shinyApp(ui, server)
```

:::{.rmdwarning}
`renderUI`: relying on it too much can create a laggy UI
:::


### Multiple controls

```{r, eval=FALSE}
library(shiny)
library(purrr)

ui <- fluidPage(
  numericInput("n", "Number of colours", value = 5, min = 1),
  uiOutput("col"),
  textOutput("palette")
)

server <- function(input, output, session) {
  col_names <- reactive(paste0("col", seq_len(input$n)))
  
  output$col <- renderUI({
    map(col_names(), ~ textInput(.x, NULL))
  })
  
  output$palette <- renderText({
    map_chr(col_names(), ~ input[[.x]] %||% "")
  })
}

shinyApp(ui, server)
```

:::{.rmdimportant}
the use of `isolate()` and `barplot()` in below example
:::

```{r, eval=FALSE}
library(shiny)
library(purrr)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      numericInput("n", "Number of colours", value = 5, min = 1),
      uiOutput("col"),
    ),
    mainPanel(
      plotOutput("plot")  
    )
  )
)

server <- function(input, output, session) {
  col_names <- reactive(paste0("col", seq_len(input$n)))
  
  output$col <- renderUI({
    map(col_names(), ~ textInput(.x, NULL, value = isolate(input[[.x]])))
  })
  
  output$plot <- renderPlot({
    cols <- map_chr(col_names(), ~ input[[.x]] %||% "")
    # convert empty inputs to transparent
    cols[cols == ""] <- NA
    
    barplot(
      rep(1, length(cols)), 
      col = cols,
      space = 0, 
      axes = FALSE
    )
  }, res = 96)
}

shinyApp(ui, server)
```

### Dynamic filtering

```{r, eval=FALSE}
library(shiny)

make_ui <- function(x, var) {
  if (is.numeric(x)) {
    rng <- range(x, na.rm = TRUE)
    sliderInput(var, var, min = rng[1], max = rng[2], value = rng)
  } else if (is.factor(x)) {
    levs <- levels(x)
    selectInput(var, var, choices = levs, selected = levs, multiple = TRUE)
  } else {
    # Not supported
    NULL
  }
}

filter_var <- function(x, val) {
  if (is.numeric(x)) {
    !is.na(x) & x >= val[1] & x <= val[2]
  } else if (is.factor(x)) {
    x %in% val
  } else {
    # No control, so don't filter
    TRUE
  }
}

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      make_ui(iris$Sepal.Length, "Sepal.Length"),
      make_ui(iris$Sepal.Width, "Sepal.Width"),
      make_ui(iris$Species, "Species")
    ),
    mainPanel(
      tableOutput("data")
    )
  )
)
server <- function(input, output, session) {
  selected <- reactive({
    filter_var(iris$Sepal.Length, input$Sepal.Length) &
      filter_var(iris$Sepal.Width, input$Sepal.Width) &
      filter_var(iris$Species, input$Species)
  })
  
  output$data <- renderTable(head(iris[selected(), ], 12))
}

shinyApp(ui, server)
```

for all columns in `iris`

```{r, eval=FALSE}
library(shiny)
library(purrr)

make_ui <- function(x, var) {
  if (is.numeric(x)) {
    rng <- range(x, na.rm = TRUE)
    sliderInput(var, var, min = rng[1], max = rng[2], value = rng)
  } else if (is.factor(x)) {
    levs <- levels(x)
    selectInput(var, var, choices = levs, selected = levs, multiple = TRUE)
  } else {
    # Not supported
    NULL
  }
}

filter_var <- function(x, val) {
  if (is.numeric(x)) {
    !is.na(x) & x >= val[1] & x <= val[2]
  } else if (is.factor(x)) {
    x %in% val
  } else {
    # No control, so don't filter
    TRUE
  }
}

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      map(names(iris), ~ make_ui(iris[[.x]], .x))
    ),
    mainPanel(
      tableOutput("data")
    )
  )
)
server <- function(input, output, session) {
  selected <- reactive({
    each_var <- map(names(iris), ~ filter_var(iris[[.x]], input[[.x]]))
    reduce(each_var, ~ .x & .y)
  })
  
  output$data <- renderTable(head(iris[selected(), ], 12))
}

shinyApp(ui, server)
```

for any dataframes in `datasets` package

```{r, eval=FALSE}
library(shiny)
library(purrr)

make_ui <- function(x, var) {
  if (is.numeric(x)) {
    rng <- range(x, na.rm = TRUE)
    sliderInput(var, var, min = rng[1], max = rng[2], value = rng)
  } else if (is.factor(x)) {
    levs <- levels(x)
    selectInput(var, var, choices = levs, selected = levs, multiple = TRUE)
  } else {
    # Not supported
    NULL
  }
}

filter_var <- function(x, val) {
  if (is.numeric(x)) {
    !is.na(x) & x >= val[1] & x <= val[2]
  } else if (is.factor(x)) {
    x %in% val
  } else {
    # No control, so don't filter
    TRUE
  }
}

dfs <- keep(ls("package:datasets"), ~ is.data.frame(get(.x, "package:datasets")))

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      selectInput("dataset", label = "Dataset", choices = dfs),
      uiOutput("filter")
    ),
    mainPanel(
      tableOutput("data")
    )
  )
)
server <- function(input, output, session) {
  data <- reactive({
    get(input$dataset, "package:datasets")
  })
  vars <- reactive(names(data()))
  
  output$filter <- renderUI(
    map(vars(), ~ make_ui(data()[[.x]], .x))
  )
  
  selected <- reactive({
    each_var <- map(vars(), ~ filter_var(data()[[.x]], input[[.x]]))
    reduce(each_var, `&`)
  })
  
  output$data <- renderTable(head(data()[selected(), ], 12))
}

shinyApp(ui, server)
```

### Exercises

#### Ex 1

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  selectInput('type', 'type', c('slider', 'numeric')),
  tabsetPanel(
    id = 'uis',
    type = 'hidden',
    tabPanel(
      'slider',
      sliderInput('n1', 'n', value = 0, min = 0, max = 100)
    ),
    tabPanel(
      'numeric',
      sliderInput('n2', 'n', value = 0, min = 0, max = 100)
    ),
  )
)

server <- function(input, output, session) {
  observeEvent(input$type, {
    updateTabsetPanel(inputId = 'uis', selected = input$type)
  })
  
  observeEvent(input$n1, {
    updateSliderInput(inputId = 'n2', value = input$n1)
  })

  observeEvent(input$n2, {
    updateSliderInput(inputId = 'n1', value = input$n2)
  })
}

shinyApp(ui, server)
```

#### Ex 2

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  actionButton("go", "Enter password"),
  textOutput("text")
)
server <- function(input, output, session) {
  observeEvent(input$go, {
    showModal(modalDialog(
      passwordInput("password", NULL, value = isolate(input$password)),
      title = "Please enter your password"
    ))
  })

  output$text <- renderText({
    if (!isTruthy(input$password)) {
      "No password"
    } else {
      "Password entered"
    }
  })
}

shinyApp(ui, server)
```

