# Why reactivity?

## Why do we need reactive programming?

`Reactive programming` is a style of programming that focuses on values that change over time, and calculations and actions that depend on those values. 

For Shiny apps to be maximally useful, we need reactive expressions and outputs to update if and only if their inputs change.

## Event-driven programming

```{r, eval=FALSE}
DynamicValue <- R6::R6Class("DynamicValue", list(
  value = NULL,
  on_update = NULL,

  get = function() self$value,

  set = function(value) {
    self$value <- value
    if (!is.null(self$on_update)) 
      self$on_update(value)
    invisible(self)
  },
  
  onUpdate = function(on_update) {
    self$on_update <- on_update
    invisible(self)
  }
))
```

```{r, eval=FALSE}
temp_c <- DynamicValue$new()
temp_c$onUpdate(function(value) {
  message("Converting") 
  temp_f <<- (value * 9 / 5) + 32
})

temp_c$set(10)

# check value of temp_f
temp_f # returns 50

# update temp_c
temp_c$set(-3)

# check value of temp_f again
temp_f # returns 26.6
```

## Reactive programming

```{r, eval=FALSE}
library(shiny)
reactiveConsole(TRUE)

# create a reactive value
temp_c <- reactiveVal(10)

# get its value
temp_c() 

# set its value
temp_c(20)

temp_c() # check its current value
```

create a reactive expression that depends on `temp_c` value:

```{r, eval=FALSE}
temp_f <- reactive({
  message("Converting")
  (temp_c() * 9 / 5) + 32
})

temp_f() # get its current value
```

:::{.rmdnote}
A reactive expression has two important properties:

+ It's lazy: it doesn't do any work until it's called.
+ It's cached: it doesn't do any work the second and subsequent times it's called because it caches the previous result.
:::

