# The reactive graph

## Introduction
Reactive inputs and expressoins are collectively called reactive producers;

Reactive expressoins and outputs are reactive consumers 

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  numericInput("a", "a", value = 10),
  numericInput("b", "b", value = 1),
  numericInput("c", "c", value = 1),
  plotOutput("x"),
  tableOutput("y"),
  textOutput("z")
)

server <- function(input, output, session) {
  rng <- reactive(input$a * 2)
  smp <- reactive(sample(rng(), input$b, replace = TRUE))
  bc <- reactive(input$b * input$c)

  output$x <- renderPlot(hist(smp()))
  output$y <- renderTable(max(smp()))
  output$z <- renderText(bc())
}

shinyApp(ui, server)
```

:::{.rmdnote}
The direction of the arrow is important: the expression records that it is used by the output; the output doesn’t record that it uses the expression. This is a subtle distinction, but its importance will become more clear when you learn about invalidation.
:::

## Dynamism

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  selectInput("choice", "A or B?", c("a", "b")),
  numericInput("a", "a", 0),
  numericInput("b", "b", 10),
  textOutput("out")
)

server <- function(input, output, session) {
  output$out <- renderText({
    if (input$choice == "a") {
      input$a
    } else {
      input$b
    }
  })
}

shinyApp(ui, server)
```
