# Reactive building blocks

## Introduction

There are 3 fundamental building blocks of reactive programming:

+ reactive values
+ reactive expressions
+ observers

## Reactive values

+ A single reactive value, created by `reactiveVal()`
+ A list of reactive values, created by `reactiveValues()`

## Reactive expressions

### Errors

:::{.rmdimportant}
Reactive expressions cache errors in exactly the same way that they cache values

```{r, eval=FALSE}
library(shiny)
reactiveConsole(TRUE)

r <- reactive(stop('Error occured at ', Sys.time(), call. = FALSE))

r()

# wait 2 seconds
Sys.sleep(2)
r()
```
:::

:::{.rmdcaution}
Errors are also treated the same way as values when it comes to the reactive graph: errors propagate through the reactive graph exactly the same way as regular values. The only difference is what happens when an error hits an output or observer:

+ An error in an output will be displayed in the app
+ An error in ab observer will cause the current session to terminate. If you don't want this to happen, you'll need to wrap the code in `try()` or `tryCatch()`

:::

## Observers and outputs

Observers and outputs differ from reactive expressions in two important ways:

+ They are eager and forgetful - they run as soon as they possibly can and they don't remember their previous action. 
+ The value returned by an observer is ignored because they are designed to work with functions called for their side-effects, like `cat()` or `write.csv()`

## Isolating code

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  numericInput("x", "x", value = 50, min = 0, max = 100),
  actionButton("capture", "capture"),
  textOutput("out")
)

server <- function(input, output, session) {
  x <- eventReactive(input$capture, input$x)

  output$out <- renderText({
    x()
  })
}

shinyApp(ui, server)
```
