# Functions

## UI Functions

Functions are a powerful tool to reduce duplication in your UI code.

```{r, eval=FALSE}
library(shiny)

sliderInput01 <- function(id) {
    sliderInput(id, label = id, min = 0, max = 1, value = 0.5, step = 0.1)
}

ui <- fluidRow(
    sliderInput01("alpha"),
    sliderInput01("beta"),
    sliderInput01("gamma"),
    sliderInput01("delta")
)
```

### functional programming

```{r}
library(purrr)
library(shiny)

vars <- c("alpha", "beta", "gamma", "delta")

sliderInput01 <- function(id) {
    sliderInput(id, label = id, min = 0, max = 1, value = 0.5, step = 0.1)
}

sliders <- map(vars, sliderInput01)

ui <- fluidRow(sliders)
```

## Server function

Whenever you have a long reactive (say > 10 lines), you should consider pulling it out into a separate function that does not use any reactivity.

+ It is much easier to debug and test your code if you can partition it so that reactivity lives inside of `server()`, and complex computation lives in your functions.
+ When looking at a reactive expression or output, there's no way to easily tell exactly what values if depends on, except by carefully reading the code block. A function definition, however, tells you exactly what the inputs are.

