# Shiny modules

The magic of modules comes because these functions are constructed in a special way that creates a "namespace".

Shiny modules have two big advantages:

+ namespacing makes it easier to understand how your app works because you can write, analyse, and test individual components in isolation.
+ modules are functions they help you reuse

## Module basics

```{r, eval=FALSE}
library(shiny)
ui <- fluidPage(
    selectInput("var", "Variable", names(mtcars)),
    numericInput("bins", "bins", 10, min = 1),
    plotOutput("hist")
)
server <- function(input, output, session) {
    data <- reactive(mtcars[[input$var]])
    output$hist <- renderPlot(
        {
            hist(data(), breaks = input$bins, main = input$var)
        },
        res = 96
    )
}
shinyApp(ui, server)
```

### module UI

```{r, eval=FALSE}
histogramUI <- function(id) {
    tagList(
        selectInput(NS(id, "var"), "Variables", choices = names(mtcars)),
        numericInput(NS(id, "bins"), "bins", value = 10, min = 1),
        plotOutput(NS(id, "hist"))
    )
}
```

### module server

```{r, eval=FALSE}
histogramServer <- function(id) {
    moduleServer(id, function(input, output, session) {
        data <- reactive(mtcars[[input$var]])
        output$hist <- renderPlot(
            {
                hist(data(), breaks = input$bins, main = input$var)
            },
            res = 96
        )
    })
}
```

updated App

```{r}
histogramApp <- function() {
    ui <- fulidPage(
        histgoramUI("hist1")
    )

    server <- function(input, output, session) {
        histogramServer("hist1")
    }

    shinyApp(ui, server)
}
```

## module server inputs and outputs

Question: how module server can take additional reactive inputs and return one or more reactive outputs?

Connecting modules together requires you to be explicit about inputs and outputs


```{r, eval=FALSE}
library(shiny)

datasetInput <- function(id, filter = NULL) {
    names <- ls("package:datasets")
    if (!is.null(filter)) {
        data <- lapply(names, get, "package:datasets")
        names <- names[vapply(data, filter, logical(1))]
    }

    selectInput(NS(id, "dataset"), "Pick a dataset", choices = names)
}

datasetServer <- function(id) {
    moduleServer(id, function(input, output, session) {
        reactive(get(input$dataset, "package:datasets"))
    })
}

datasetApp <- function(filter = NULL) {
    ui <- fluidPage(
        datasetInput("dataset", filter = filter),
        tableOutput("data")
    )
    server <- function(input, output, session) {
        data <- datasetServer("dataset")
        output$data <- renderTable(head(data()))
    }
    shinyApp(ui, server)
}
```

### Add variable selection

```{r, eval=FALSE}
library(shiny)

datasetInput <- function(id, filter = NULL) {
    names <- ls("package:datasets")
    if (!is.null(filter)) {
        data <- lapply(names, get, "package:datasets")
        names <- names[vapply(data, filter, logical(1))]
    }

    selectInput(NS(id, "dataset"), "Pick a dataset", choices = names)
}

datasetServer <- function(id) {
    moduleServer(id, function(input, output, session) {
        reactive(get(input$dataset, "package:datasets"))
    })
}

# select variable UI
selectVarInput <- function(id) {
    selectInput(NS(id, "var"), "Variable", choices = NULL)
}

# select variable server: helper function
find_vars <- function(data, filter) {
    names(data)[vapply(data, filter, logical(1))]
}

# select variable server
selectVarServer <- function(id, data, filter = is.numeric) {
    moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
            updateSelectInput(session, "var", choices = find_vars(data(), filter))
        })

        reactive(data()[[input$var]])
    })
}

# the updated App: with variable selection included
selectVarApp <- function(filter = is.numeric) {
    ui <- fluidPage(
        datasetInput("data", is.data.frame),
        selectVarInput("var"),
        verbatimTextOutput("out")
    )

    server <- function(input, output, session) {
        data <- datasetServer("data")
        var <- selectVarServer("var", data, filter = filter)

        output$out <- renderPrint(var())
    }

    shinyApp(ui, server)
}

# selectVarApp()
```

## Server inputs

When designing a module server, we need to think about who is going to provide the value for each argument:

+ is it fixed and constant over the life-time of the app?
+ or is it reactive, changing as the user interacts with the app?

Once we've made this decision, it's good practice to check that each input to the server module is either reactive or constant.

:::{.rmdimportant}
`input$var` is not a reactive, so whenever you pass an input value into a module, you’ll need to wrap it in a `reactive()`
:::

### updated version of `selectVarServer()`

```{r, eval=FALSE}
selectVarServer <- function(id, data, filter = is.numeric) {
    stopifnot(is.reactive(data))
    stopifnot(!is.reactive(filter))

    moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
            updateSelectInput(session, "var", choices = find_vars(data(), filter))
        })

        reactive(data()[[input$var]])
    })
}
```

### Modules inside of modules

```{r, eval=FALSE}
library(shiny)

datasetInput <- function(id, filter = NULL) {
    names <- ls("package:datasets")
    if (!is.null(filter)) {
        data <- lapply(names, get, "package:datasets")
        names <- names[vapply(data, filter, logical(1))]
    }

    selectInput(NS(id, "dataset"), "Pick a dataset", choices = names)
}

datasetServer <- function(id) {
    moduleServer(id, function(input, output, session) {
        reactive(get(input$dataset, "package:datasets"))
    })
}

# select variable UI
selectVarInput <- function(id) {
    selectInput(NS(id, "var"), "Variable", choices = NULL)
}

# select variable server: helper function
find_vars <- function(data, filter) {
    names(data)[vapply(data, filter, logical(1))]
}

# select variable server
selectVarServer <- function(id, data, filter = is.numeric) {
    stopifnot(is.reactive(data))
    stopifnot(!is.reactive(filter))

    moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
            updateSelectInput(session, "var", choices = find_vars(data(), filter))
        })

        reactive(data()[[input$var]])
    })
}

selectDataVarUI <- function(id) {
    tagList(
        datasetInput(NS(id, "data"), filter = is.data.frame),
        selectVarInput(NS(id, "var"))
    )
}

selectDataVarServer <- function(id, filter = is.numeric) {
    moduleServer(id, function(input, output, session) {
        data <- datasetServer("data")
        var <- selectVarServer("var", data, filter = filter)

        var
    })
}

selectDataVarApp <- function(filter = is.numeric) {
    ui <- fluidPage(
        sidebarLayout(
            sidebarPanel(selectDataVarUI("var")),
            mainPanel(verbatimTextOutput("out"))
        )
    )

    server <- function(input, output, session) {
        var <- selectDataVarServer("var", filter)

        output$out <- renderPrint(var(), width = 40)
    }

    shinyApp(ui, server)
}

# selectDataVarApp()
```



### histgoram example

:::{.rmdnote}
The key challenge of creating modules is creating functions that are flexible enough to be used in multiple places, but simple enough that they can easily be understood.
:::

```{r, eval=FALSE}
library(shiny)

datasetInput <- function(id, filter = NULL) {
    names <- ls("package:datasets")
    if (!is.null(filter)) {
        data <- lapply(names, get, "package:datasets")
        names <- names[vapply(data, filter, logical(1))]
    }

    selectInput(NS(id, "dataset"), "Pick a dataset", choices = names)
}

datasetServer <- function(id) {
    moduleServer(id, function(input, output, session) {
        reactive(get(input$dataset, "package:datasets"))
    })
}

# select variable UI
selectVarInput <- function(id) {
    selectInput(NS(id, "var"), "Variable", choices = NULL)
}

# select variable server: helper function
find_vars <- function(data, filter) {
    names(data)[vapply(data, filter, logical(1))]
}

# select variable server
selectVarServer <- function(id, data, filter = is.numeric) {
    stopifnot(is.reactive(data))
    stopifnot(!is.reactive(filter))

    moduleServer(id, function(input, output, session) {
        observeEvent(data(), {
            updateSelectInput(session, "var", choices = find_vars(data(), filter))
        })

        reactive(data()[[input$var]])
    })
}

# histogram example

histogramOutput <- function(id) {
    tagList(
        numericInput(NS(id, "bins"), "bins", 10, min = 1, step = 1),
        plotOutput(NS(id, "hist"))
    )
}

# x: variable to plot
# tilte: title for the histogram
histogramServer <- function(id, x, title = reactive("Histogram")) {
    stopifnot(is.reactive(x))
    stopifnot(is.reactive(title))

    moduleServer(id, function(input, output, session) {
        output$hist <- renderPlot(
            {
                main <- paste0(title(), " [", input$bins, "]")
                hist(x(), breaks = input$bins, main = main)
            },
            res = 96
        )
    })
}

histogramApp <- function() {
    ui <- fluidPage(
        sidebarLayout(
            sidebarPanel(
                datasetInput("data", is.data.frame),
                selectVarInput("var")
            ),
            mainPanel(
                histogramOutput("hist")
            )
        ),
    )

    server <- function(input, output, session) {
        data <- datasetServer("data")
        x <- selectVarServer("var", data)

        histogramServer("hist", x)
    }

    shinyApp(ui, server)
}

histogramApp()
```

:::{.rmdimportant}
When passing an argument to a module, you want the module to react to the value changing which means that you have to pass the reactive, not its **current value**.
:::

## limited selection + other

```{r, eval=FALSE}
library(shiny)

radioExtraUI <- function(id, label, choices, selected = NULL, placeholder = "Other") {
    other <- textInput(NS(id, "other"), label = NULL, placeholder = placeholder)

    names <- if (is.null(names(choices))) choices else names(choices)
    values <- unname(choices)

    radioButtons(NS(id, "primary"),
        label = label,
        choiceValues = c(names, "other"),
        choiceNames = c(as.list(values), list(other)),
        selected = selected
    )
}

radioExtraServer <- function(id) {
    moduleServer(id, function(input, output, session) {
        observeEvent(input$other, ignoreInit = TRUE, {
            updateRadioButtons(session, "primary", selected = "other")
        })

        reactive({
            if (input$primary == "other") {
                input$other
            } else {
                input$primary
            }
        })
    })
}

radioExtraApp <- function(...) {
    ui <- fluidPage(
        radioExtraUI("extra", ...),
        textOutput("value")
    )
    server <- function(input, output, server) {
        extra <- radioExtraServer("extra")
        output$value <- renderText(paste0("Selected: ", extra()))
    }

    shinyApp(ui, server)
}

radioExtraApp(
    label = "How do you usually read csv files?",
    choices = c("read.csv()", "readr::read_csv()", "data.table::fread()"),
    selected = "read.csv()"
)
```


## Wizard

```{r, eval=FALSE}
library(shiny)

nextPage <- function(id, i) {
    actionButton(NS(id, paste0("go_", i, "_", i + 1)), "next")
}

prevPage <- function(id, i) {
    actionButton(NS(id, paste0("go_", i, "_", i - 1)), "prev")
}

wrapPage <- function(title, page, button_left = NULL, button_right = NULL) {
    tabPanel(
        title = title,
        fluidRow(
            column(12, page)
        ),
        fluidRow(
            column(6, button_left),
            column(6, button_right)
        )
    )
}

wizardUI <- function(id, pages, doneButton = NULL) {
    stopifnot(is.list(pages))
    n <- length(pages)

    wrapped <- vector("list", n)

    for (i in seq_along(pages)) {
        # first page only has next; last page only has prev + done
        lhs <- if (i > 1) prevPage(id, i)
        rhs <- if (i < n) nextPage(id, i) else doneButton

        wrapped[[i]] <- wrapPage(paste0("page_", i), pages[[i]], lhs, rhs)
    }

    # create tabsetPanel
    wrapped$id <- NS(id, "wizard")
    wrapped$type <- "hidden"
    do.call("tabsetPanel", wrapped)
}

wizardServer <- function(id, n) {
    moduleServer(id, function(input, output, session) {
        changePage <- function(from, to) {
            observeEvent(input[[paste0("go_", from, "_", to)]], {
                updateTabsetPanel(session, "wizard", selected = paste0("page_", to))
            })
        }

        ids <- seq_len(n)

        lapply(ids[-1], function(i) changePage(i, i - 1))
        lapply(ids[-n], function(i) changePage(i, i + 1))
    })
}

wizardApp <- function(...) {
    pages <- list(...)

    ui <- fluidPage(
        wizardUI("whiz", pages)
    )

    server <- function(input, output, session) {
        wizardServer("whiz", length(pages))
    }

    shinyApp(ui, server)
}

page1 <- tagList(
    textInput("name", "What's your name?")
)
page2 <- tagList(
    numericInput("age", "How old are you?", 20)
)
page3 <- tagList(
    "Is this data correct?",
    verbatimTextOutput("info")
)

# call 1
wizardApp(page1, page2, page3)

ui <- fluidPage(
    wizardUI(
        id = "demographics",
        pages = list(page1, page2, page3),
        doneButton = actionButton("done", "Submit")
    )
)

server <- function(input, output, session) {
    wizardServer("demographics", 3)

    observeEvent(input$done, showModal(
        modalDialog("Thank you!", footer = NULL)
    ))

    output$info <- renderText(paste0(
        "Age: ", input$age, "\n",
        "Name: ", input$name, "\n"
    ))
}
# call 2
shinyApp(ui, server)
```

## Dynamic UI

```{r, eval=FALSE}
library(shiny)
library(purrr)

filterUI <- function(id) {
    uiOutput(NS(id, "controls"))
}

make_ui <- function(x, id, var) {
    if (is.numeric(x)) {
        rng <- range(x, na.rm = TRUE)
        sliderInput(id, var, min = rng[1], max = rng[2], value = rng)
    } else if (is.factor(x)) {
        levs <- levels(x)
        selectInput(id, var, choices = levs, selected = levs, multiple = TRUE)
    } else {
        NULL
    }
}

filter_var <- function(x, val) {
    if (is.numeric(x)) {
        !is.na(x) & x >= val[1] & x <= val[2]
    } else if (is.factor(x)) {
        x %in% val
    } else {
        # No control, so don't filter
        TRUE
    }
}

filterServer <- function(id, df) {
    stopifnot(is.reactive(df))

    moduleServer(id, function(input, output, session) {
        vars <- reactive(names(df()))

        output$controls <- renderUI({
            map(vars(), function(var) make_ui(df()[[var]], NS(id, var), var))
        })

        reactive({
            each_var <- map(vars(), function(var) filter_var(df()[[var]], input[[var]]))
            reduce(each_var, `&`)
        })
    })
}

datasetInput <- function(id, filter = NULL) {
    names <- ls("package:datasets")
    if (!is.null(filter)) {
        data <- lapply(names, get, "package:datasets")
        names <- names[vapply(data, filter, logical(1))]
    }

    selectInput(NS(id, "dataset"), "Pick a dataset", choices = names)
}

datasetServer <- function(id) {
    moduleServer(id, function(input, output, session) {
        reactive(get(input$dataset, "package:datasets"))
    })
}

filterApp <- function() {
    ui <- fluidPage(
        sidebarLayout(
            sidebarPanel(
                datasetInput("data", is.data.frame),
                textOutput("n"),
                filterUI("filter")
            ),
            mainPanel(
                tableOutput("table")
            )
        )
    )

    server <- function(input, output, session) {
        df <- datasetServer("data")
        filter <- filterServer("filter", df)

        output$table <- renderTable({
            df()[filter(), , drop = FALSE]
        })

        output$n <- renderText({
            paste0(sum(filter()), " rows")
        })
    }

    shinyApp(ui, server)
}

filterApp()
```